---
title: "Final"
author: "Sri Seshadri"
date: "5/20/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(magrittr)
library(tidyverse)
library(skimr)
library(forecast)
library(lubridate)
library(xts)
library(astsa)
library(caret)
library(dplyr)
```

#1. Introduction


#2. Executive Summary

#3. About the data

```{r, warning = F}
train <- read_csv("Training.csv")
test <- read_csv("DengAI_Predicting_Disease_Spread_-_Test_Data_Features.csv")

skim_with(numeric= list(hist= NULL),integer = list(hist = NULL))
pander::pander(skim(train))
# Show missing values 
```

```{r, fig.cap="Missing data map"}
Amelia::missmap(train,y.cex = 0.5,x.cex = 0.5)
```

# 4. Data Preparation

Split the data into the 2 regions - based on a paper that different regions had different Dengue reports based on localized conditions.
Impute data based on kNN method - chicken or the egg problem... impute data before feature selection or feature selection before imputing?
Verification of data imputing covered in the next section.

```{r}

# Split data by city

trainSJ <- train %>% dplyr::filter(city == "sj")
trainIQ <- train %>% dplyr::filter(city == "iq")

trainImputedSj <- VIM::kNN(trainSJ[, -5])
trainImputedIq <- VIM::kNN(trainIQ[, -5])

trainImputedSj %<>% mutate(week_start_date = lubridate::mdy(week_start_date),
                           total_cases = trainSJ$total_cases)
trainImputedIq %<>% mutate(week_start_date = lubridate::mdy(week_start_date),
                           total_cases = trainIQ$total_cases)

```

# 5. Exploratory Data Analysis (EDA)

Reported cases - non stationary, seasonal series

```{r}
par(mfrow = c(2,1))
tsplot(trainImputedSj$total_cases~trainImputedSj$week_start_date, main = "Dengue reported cases in San Juan", ylab = "Total cases")
tsplot(trainImputedIq$total_cases~trainImputedIq$week_start_date, main = "Dengue reported cases in Iquitos", ylab = "Total cases")
```

## verify imputation

```{r}
plot(trainImputedSj$ndvi_ne~trainImputedSj$week_start_date, col = ifelse(trainImputedSj$ndvi_ne_imp == T,"red", "black"),type = "b")
plot(trainImputedIq$ndvi_ne~trainImputedIq$week_start_date, col = ifelse(trainImputedIq$ndvi_ne_imp == T,"red", "black"), type = "b")

plot(trainImputedSj$ndvi_nw~trainImputedSj$week_start_date, col = ifelse(trainImputedSj$ndvi_nw_imp == T,"red", "black"),type = "b")
plot(trainImputedIq$ndvi_nw~trainImputedIq$week_start_date, col = ifelse(trainImputedIq$ndvi_nw_imp == T,"red", "black"), type = "b")

plot(trainImputedSj$station_avg_temp_c~trainImputedSj$week_start_date, col = ifelse(trainImputedSj$station_avg_temp_c_imp == T,"red", "black"),type = "b")
plot(trainImputedIq$station_avg_temp_c~trainImputedIq$week_start_date, col = ifelse(trainImputedIq$station_avg_temp_c_imp == T,"red", "black"),type = "b",lty = 1,lwd = 0.5, cex = 0.5)

```




## 

```{r}
cormatSj <-  trainImputedSj %>% 
  select_if(is.numeric) %>% 
  cor(.)

corrplot::corrplot(cormatSj,method = 'ellipse',tl.cex = 0.6,order = "hclust")


cormatIq <- trainImputedIq %>% 
  select_if(is.numeric) %>% 
  filter(year > 2001) %>% 
  cor(.)

corrplot::corrplot(cormatIq, method = 'ellipse',tl.cex = 0.6, order = 'hclust')

HighCorIq <- findCorrelation(cormatIq, names = T)
HighCorSj <- findCorrelation(cormatSj, names = T)

HighCorIq
HighCorSj
```

```{r}
tsTotalcasesSj <- ts(trainImputedSj$total_cases, start = c(1990,18), frequency = 52)
ggseasonplot(tsTotalcasesSj)  



ggseasonplot(ts(trainImputedSj$reanalysis_relative_humidity_percent,start = c(1990,18), frequency = 52)) + theme_classic() +theme(legend.position = "none",plot.title = element_text(hjust = 0.5),axis.text.x = element_text(size = 5)) + ggtitle("Reanalysis_relative_humidity_%")

ggseasonplot(ts(trainImputedSj$reanalysis_max_air_temp_k, start = c(1990,18), frequency = 52)) +  theme_classic() + theme(legend.position = "none",plot.title = element_text(hjust = 0.5),axis.text.x = element_text(size = 5)) + ggtitle("Reanalysis_max_air_temp_K")


tsTotalcasesIq <- ts(trainImputedIq$total_cases, start = c(2000,26), frequency = 52)
ggseasonplot(tsTotalcasesIq)  

ggseasonplot(ts(trainImputedIq$reanalysis_relative_humidity_percent,start = c(2000,26), frequency = 52)) + theme_classic() +theme(legend.position = "none",plot.title = element_text(hjust = 0.5),axis.text.x = element_text(size = 5)) + ggtitle("Reanalysis_relative_humidity_% - Iq")

ggseasonplot(ts(trainImputedIq$reanalysis_max_air_temp_k, start = c(2000,26), frequency = 52)) +  theme_classic() + theme(legend.position = "none",plot.title = element_text(hjust = 0.5),axis.text.x = element_text(size = 5)) + ggtitle("Reanalysis_max_air_temp_K - IQ")
```


# 6. Feature Selection
 
1. Recursive feature elimination 

```{r}
# Remove highCor variables
trainImputedSj %<>% 
  select(-city,-Year.Week) %>%  
  select(-one_of(!!HighCorSj)) %>% 
  select(-ends_with("imp"))

# make data a time series object
tsTrainImputedSj <-  xts::xts(trainImputedSj[,-1:-3],order.by = trainImputedSj$week_start_date, frequency = 52)

#ccf2(trainImputedSj$total_cases,trainImputedSj$reanalysis_specific_humidity_g_per_kg,max.lag = 15)

# Removing 2001 data for analysis and feature selection. Assuming data was not properly collected during those times
trainImputedIq %<>% 
  filter(year > 2001) %>% 
  select(-city,-Year.Week) %>% 
  select(-one_of(!!HighCorIq)) %>% 
  select(-ends_with("imp"))

tsTrainImputedIq <-  xts::xts(trainImputedIq[,-1:-3],order.by = trainImputedIq$week_start_date, frequency = 52)
```

## Random forest

```{r}

# Fearure selection by removing time as a variable.

newRF <- rfFuncs
varSeq <- seq(1,15,by = 2)
ctrl <- rfeControl(method = "repeatedcv",
                   repeats = 5,
                   verbose = F,
                   functions = rfFuncs,
                   allowParallel = T)
rfRFE <- rfe(x = trainImputedSj[,c(-1:-3,-19)],
             y = trainImputedSj$total_cases,
             sizes = varSeq,
             metric = "RMSE",
             rfeControl = ctrl,
             ntree = 500)
             #ntree = 1000)


predictors(rfRFE)

# Iq

rfRFEiq <- rfe(x = trainImputedIq[,c(-1:-3,-21)],
             y = trainImputedIq$total_cases,
             sizes = varSeq,
             metric = "RMSE",
             rfeControl = ctrl,
             ntree = 500)
             #ntree = 1000)


predictors(rfRFEiq)


```

### random forest after removing the effect of time

```{r}
# remove 1996 data for feature selection
auto.arima(tsTrainImputedSj$total_cases)
auto.arima(trainImputedSj$total_cases[trainImputedSj$year >= 1996],stepwise = F)
astsa::acf2(trainImputedSj$total_cases[trainImputedSj$year >= 1996])
fit <- astsa::sarima(trainImputedSj$total_cases[trainImputedSj$year >= 1996],p = 1,d = 1, q = 1)
Box.test(fit$fit$residuals)
auto.arima(trainImputedSj$total_cases[trainImputedSj$year >= 1996],stepwise = F)

```

```{r}
rfRFE_timeEffect <- rfe(x = trainImputedSj[trainImputedSj$year >= 1996,c(-1:-3,-19)],
             y = fit$fit$residuals,
             sizes = varSeq,
             metric = "RMSE",
             rfeControl = ctrl,
             ntree = 500)

rfRFE_timeEffect$optVariables
```

```{r}
x <- model.matrix(total_cases ~ ., data = trainImputedSj[,-1:-3])[,-1]
cv.out <- glmnet::cv.glmnet(x,y = trainImputedSj$total_cases, family = "poisson", alpha = 1)
plot(cv.out)
out <- glmnet::glmnet(x,trainImputedSj$total_cases, family = 'poisson',lambda = cv.out$lambda.min, alpha = 1)
topPredLasso_PresentTime <- abs(predict(out,type = "coefficients",s=cv.out$lambda.min))

topPredL_presentTime <- data.frame(Predictors = topPredLasso_PresentTime@Dimnames[[1]][topPredLasso_PresentTime@i+1],
Coefs = topPredLasso_PresentTime@x) 

topPredL_presentTime %<>% arrange(desc(Coefs)) %>% filter(!Predictors == "(Intercept)")
```

 
```{r}
Box.test(fit$fit$residuals)
cv.out <- glmnet::cv.glmnet(x = model.matrix(total_cases ~ ., trainImputedSj[trainImputedSj$year >= 1996,c(-1:-3)])[,-1],y = as.numeric(fit$fit$residuals), alpha = 1)
out <- glmnet::glmnet(x,trainImputedSj$total_cases,lambda = cv.out$lambda.min, alpha = 1)
topPredLasso <- abs(predict(out,type = "coefficients",s=cv.out$lambda.min))
topPredLasso
topPredL <- data.frame(Predictors = topPredLasso@Dimnames[[1]][topPredLasso@i+1],
Coefs = topPredLasso@x) 

topPredL %<>% arrange(desc(Coefs)) %>% filter(!Predictors == "(Intercept)")
```


# Are the predictors stationary?

```{r}

```

